---
title: "Experimentation Log"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Declare Variables and Import Data

It will be useful to store variable like this, so when we try to generalize it into an app, it is more straightforward.
```{r}
number_of_variables <- 3
form <- Age~Weight+Oxygen
n <- 200

data <- read.csv(file="~/Desktop/5763GroupProject/data/fitness.csv", header=TRUE, sep=",")
```

## Get quantiles

```{r}

get_quantiles <- function(coeff, num_var){
  for(i in 1:num_var){
    #get quantiles
    #how does one accesss the index of a quantile
    print(quantile(coeff[,i], probs = c(0.025,0.975)))
  }
}

```

## Original Boot Data Function (Slightly Modified)

```{r}
baselineBootstrap <- function(inputData, num_var,formula, nBoots){
  
  for(i in 1:nBoots){
    #randomly sample data
    bootData <- inputData[sample(1:nrow(inputData), nrow(inputData), replace = T),]
    bootLM <- lm(formula, data = bootData)
    # store the coefs
    #for optimization put 
    if(i == 1){
      bootResults <- matrix(coef(bootLM), ncol = num_var)
    } else {
      bootResults<- rbind(bootResults, matrix(coef(bootLM), ncol = num_var))
    }
  } # end of i loop
  return(bootResults)
}

```

Get the coefficients for a linear model trained on the declared variables above and print quantiles

```{r}
set.seed(9)
coefficients1 <- baselineBootstrap(inputData = data,
                                   num_var = number_of_variables,
                                   formula = form,
                                   nBoots = n)

get_quantiles(coefficients1,number_of_variables)
```

## SpeededUp 1 

What I tried to improve:

Instantiate matrix with all zeros instead of rbind hell. 

An experiment will need to occur with instantiating a nan matrix
- (10/15) Result is failure. Instantiating with NaNs slows performance on all metrics but Median
```{r}
speedyBoot <- function(inputData, num_var,formula, nBoots){
  mat <- matrix(0L, nrow = nBoots, ncol = num_var)
  for(i in 1:nBoots){
    bootData <- inputData[sample(1:nrow(inputData), nrow(inputData), replace = T),]
    bootLM <- lm(formula, data = bootData)
    # store the coefs
    mat[i,] <- coef(bootLM)
  } # end of i loop
  return(mat)
}

```

Get coefficients and assess quantiles

```{r}
set.seed(9)

coefficients2 <- speedyBoot(inputData = data,
                                   num_var = number_of_variables,
                                   formula = form,
                                   nBoots = n)
get_quantiles(coefficients2, number_of_variables)

```


## Comparison of algorithms 

One wants n large so that the gap in performance is larger and more evident. If n is too small, then the performance improvement will be negligible.

```{r}

set.seed(9)

library(microbenchmark)
microbenchmark(
  coefficients1 <- baselineBootstrap(inputData = data,
                                   num_var = number_of_variables,
                                   formula = form,
                                   nBoots = n),
  coefficients2 <- speedyBoot(inputData = data,
                                   num_var = number_of_variables,
                                   formula = form,
                                   nBoots = n)
  
  )  
```
There is improvement. However the difference is so nominal that it could be completely stochastic. 

## Speedy 2

Try to improve random sampling bottleneck by removing list enumeration in sample
```{r}

speedyBoot2 <- function(inputData, num_var,formula, nBoots){
  mat <- matrix(0L, nrow = nBoots, ncol = num_var)
  for(i in 1:nBoots){
    bootData <- inputData[sample(nrow(inputData), nrow(inputData), replace = T),]
    bootLM <- lm(formula, data = bootData)
    # store the coefs
    mat[i,] <- coef(bootLM)
  } # end of i loop
  return(mat)
}

set.seed(9)

coefficients3 <- speedyBoot2(inputData = data,
                                   num_var = number_of_variables,
                                   formula = form,
                                   nBoots = n)
get_quantiles(coefficients3, number_of_variables)
```

```{r}
set.seed(8)
n<- 1000
microbenchmark(
  coefficients1 <- baselineBootstrap(inputData = data,
                                   num_var = number_of_variables,
                                   formula = form,
                                   nBoots = n),
  coefficients3 <- speedyBoot2(inputData = data,
                                   num_var = number_of_variables,
                                   formula = form,
                                   nBoots = n)
  
  )  
```

## experiment 4

```{r}
speedy4 <- function(inputData, num_var,formula){
    bootLM <- lm(formula, data = inputData[sample(nrow(inputData), nrow(inputData), replace = T),])
     return(coef(bootLM))
}

```


make function

```{r}
library(foreach)
library(doParallel)
n<-200
set.seed(9)

cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)
clusterSetRNGStream(cl=cl,9)

bootCoefs <- foreach(i = 1:n, .combine =rbind) %dopar% speedy4(inputData = data,
                                   num_var = number_of_variables,
                                   formula = form)


get_quantiles(bootCoefs, number_of_variables)
stopCluster(cl)

```

Check speed
```{r}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)
clusterSetRNGStream(cl=cl,9)
n<-1000

microbenchmark(
  coefficients1 <- baselineBootstrap(inputData = data,
                                   num_var = number_of_variables,
                                   formula = form,
                                   nBoots = n),
  
  bootCoefs <- foreach(i = 1:n, .combine = rbind) %dopar% speedy4(
                                    inputData = data,
                                   num_var = number_of_variables,
                                   formula = form)

  )  
stopCluster(cl)

```
